Testing Quasar Unification with Clumpy Wind Models

### 2015-11-10 paper.tex
rm 0946

Perhaps the clearest evidence of outflows in AGN is  
the blueshifted ($\sim 0.1c$) broad absorption lines (BALs) in the 
ultraviolet seen in approximately $20\%$ of quasars
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within the ADW paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
Polarisation studies expect the wind to be roughly equatorial
\citep{goodrich1995, cohen1995}, although there is also evidence
for polar BAL outflows  in radio-loud (RL) sources \citep{zhou2006}.
% although this geometry is contradicted
% by more recent studies of radio spectral indices \citep{dipompeo2012}.
ADWs offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.

Within the ADW  unification framework, the broad-line region (BLR) can 
correspond either to the dense wind base or dense clumps embedded
in the outflows. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

@@

%{\bf ksl: I don't see what this paragraph has to do with our problem, which we need to clearly explain} Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics,
Despite the clear importance of ADWs in shaping quasar and AGN spectra,  
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
% The presence of line-locked features \citep{bowler2014} 
% and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
% also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
% in the spectra of some BALQSOs also gives clearer evidence that line-driving is
% at least partially contributing to the acceleration of the wind.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
% Hydrodynamic simulations of line-driven disc winds also found a shielding region
% was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
% However, \cite{sim2010_hydro} and \cite{H14} showed that including multiple scattering means the ionizing radiation field could still reach the previously shielded regions 
% in those particular models.
An additional or alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

@@

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BAL QSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.   The results have encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BAL QSOs, as long as the luminosity of the central engine was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

Here we attempt to address both of these issues, by introducing clumping into our model and a more complete treatment of H and He into our radiative transfer calculations.   The remainder of this paper is organized as follows:
% There has been some success using simple kinematic prescriptions for biconical disc winds to model  AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  H13 successfully produced a benchmark model for BALQSOs. However, the model had two key drawbacks. First, an unrealistically low X-ray luminosity was required in order to prevent over-ionization of the outflow. Second, the model failed to produce the strong emission lines  required at low inclinations in a unified model. In this paper, we attempt to address these issues, and test the disc wind  unification model using Monte Carlo radiative transfer (MCRT) and photoionization  calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. In section 3, we outline the model, including 
a description of our clumping implementation. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

For this study, we use the MCRT code \py\ we have developed to carry out our 
radiative transfer and photoionization calculations in non-local thermodynamic 
equilibrium (non-LTE) in a variety of disk-wind systems. 
The code has been used to model accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects \citep{simmacro2005}
and quasars (H13).\nocite{noebauer, M15, LK02}  
The basic premise involves generating photons from an accretion disc
and central object, and computing their radiative transfer through an outflow.
This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry.
First, the code gradually converges on a global temperature and ionization structure.
Once these ``ionization cycles'' are complete, then a series of further iterations
are completed to obtain a detailed synthetic spectrum.
LK02 provides the basic description of the original code; improvements, 
see especially H14 and M15, have been described in subsequent reports.  
%A full description of the code will be published elsewhere; 
Therefore we focus here on the changes made to improve the ionization calculation 
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/link.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:wind}
\end{figure*} %fullpage

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Next-generation Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{ergs~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f$ 	&	 $0.01^*$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters used in the model.}
\label{wind_param}
\end{table}

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
density squared. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/0946.png}
\caption
{
Synthetic spectra at $72^\circ$ compared to an HST STIS spectrum
of PG0946+301 (Arav et al. 2000). The spectrum is scaled to the flux at 1300\AA\
for direct comparison. 
}
\label{fig:stis}
\end{figure*} %fullpage

@@

Nonblack saturation is observed in the absorption troughs of BALQSOs \citep{arav1999a,arav1999b}.
This can be caused either by partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not cospatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are readily apparent. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To do this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{surlan2012,hamann2008}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered contribution is obscured by the saturated troughs.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.

@@

We find significant collisionally excited line emission emerges
at low inclinations in the synthetic spectra, particular in the \civ line.
The improved treatment of recombination also results in a strong \la\ line. 
In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. 
The spectra do not contain the strong C~\textsc{iii}]~1909\AA\, 
line seen in the quasar composite spectra. 
This is because we do not yet treat C as a full macro-atom with a 
full collisional rates between forbidden
or semi-forbidden transitions, as would be required.
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
outer portion of our wind. 
We therefore expect a model with these 
parameters to produce a C~\textsc{iii}]~1909\AA\ line
with a proper treament.

>>>>

Perhaps the clearest evidence of outflows in AGN is  
the blueshifted ($\sim 0.1c$) broad absorption lines (BALs) in the 
ultraviolet seen in approximately $20\%$ of quasars
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind. 
According to this paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
Polarisation studies expect the wind to be roughly equatorial
\citep{goodrich1995, cohen1995}, although there is also evidence
for polar BAL outflows  in radio-loud (RL) sources \citep{zhou2006}.
% although this geometry is contradicted
% by more recent studies of radio spectral indices \citep{dipompeo2012}.

Due to their ubiquitous nature,
disc winds offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this unification framework, the broad-line region (BLR) can 
correspond either to the dense wind base or clumps embedded
in the outflow. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

@@

%{\bf ksl: I don't see what this paragraph has to do with our problem, which we need to clearly explain} Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics,
Despite the clear importance of disc winds in shaping quasar and AGN spectra,  
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
% The presence of line-locked features \citep{bowler2014} 
% and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
% also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
% in the spectra of some BALQSOs also gives clearer evidence that line-driving is
% at least partially contributing to the acceleration of the wind.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
% Hydrodynamic simulations of line-driven disc winds also found a shielding region
% was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
% However, \cite{sim2010_hydro} and \cite{H14} showed that including multiple scattering means the ionizing radiation field could still reach the previously shielded regions 
% in those particular models.
An additional or alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

@@

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BAL QSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer (MCRT) code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  The results have encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BAL QSOs, as long as the luminosity of the central engine was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

Here we attempt to address both of these issues, by introducing clumping into our model and a more complete treatment of H and He into our radiative transfer calculations.   The remainder of this paper is organized as follows:
% There has been some success using simple kinematic prescriptions for biconical disc winds to model  AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  H13 successfully produced a benchmark model for BALQSOs. However, the model had two key drawbacks. First, an unrealistically low X-ray luminosity was required in order to prevent over-ionization of the outflow. Second, the model failed to produce the strong emission lines  required at low inclinations in a unified model. In this paper, we attempt to address these issues, and test the disc wind  unification model using Monte Carlo radiative transfer (MCRT) and photoionization  calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. We then outline the model in section 3, including 
a description of our clumping implementation and success criteria. 
Section 4 contains the results from a clumped model, which we discuss results
in comparison to observational data. Finally, we summarise our findings in section 5.

@@

For this study, we use the MCRT code \py\ we have developed to carry out our 
radiative transfer and photoionization simulations in non-local thermodynamic
equilibrium (non-LTE). The code can be used to model a variety of
disc-wind systems; it has been used with application to accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The basic premise involves generating photons from an accretion disc
and central object, and computing their radiative transfer through an outflow.
This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry.
First, the code gradually converges on the global temperature and ionization structure.
Once these ``ionization cycles'' are complete, then a series of further iterations
allow one to obtain a detailed synthetic spectrum.
LK02 provides the basic description of the original code; improvements, 
see especially \cite{simmacro2005}, H13 and M15, have been 
described in subsequent reports.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made to improve the ionization calculation 
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section, $n_e$
is the electron density and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative and 
collisional (three-body) recombination, respectively. 
$T_e$ is the electron temperature.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$. 
We stress that this approximation is not required for ions
treated as macro-atoms.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/link8.png}
\caption
{
Contour plots showing the logarithm of some important 
physical properties of the outflow. Symbols are defined in the text.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
The line luminosities represent the luminosity of photons
escaping the Sobolev region for each line. These photons do not
necessarily escape to infinity.
}
\label{fig:wind}
\end{figure*} %fullpage

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Next-generation Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{ergs~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters used in the model.}
\label{wind_param}
\end{table}

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
This is illustrated in figure~\ref{fig:wind} 
by the panels showing the ion fraction $F=n_j/n_{tot}$ of some important ions.
By clumping the wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
density squared. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. The line luminosities 
shown in the figure correspond to the luminosity in $\LUM$ of photons
escaping the Sobolev region for each line. 
This is equivalent to $\beta_{ul} n_u A_{ul}$, where the three quantities
represent the Sobolev escape probability, upper level number density and 
Einstein A coefficient for the line. 
As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

% \begin{figure*} %fullpage
% \centering
% \includegraphics[width=1.0\textwidth]{figures/0946.png}
% \caption
% {
% Synthetic spectra at $72^\circ$ compared to an HST STIS spectrum
% of PG0946+301 (Arav et al. 2000). The spectrum is scaled to the flux at 1300\AA\
% for direct comparison. 
% }
% \label{fig:stis}
% \end{figure*} %fullpage

@@

Nonblack saturation is observed in the absorption troughs of BALQSOs \citep{arav1999a,arav1999b}.
This can be caused either by partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not cospatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are readily apparent. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To do this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{surlan2012,hamann2008}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.

@@

We find significant collisionally excited line emission emerges
at low inclinations in the synthetic spectra, particular in the \civ\ line.
The improved treatment of recombination also results in a strong \la\ line. 
In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. 
The spectra do not contain the strong C~\textsc{iii}]~1909\AA\, 
line seen in the quasar composite spectra. 
This is because we do not yet treat C as a full macro-atom with a 
full collisional rates between forbidden
or semi-forbidden transitions, as would be required.
The critical density of the C~\textsc{iii}]~1909\AA\, line 
is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
outer portion of our wind. 
We therefore expect a model with these 
parameters to produce a C~\textsc{iii}]~1909\AA\ line
with a proper treament.
